### 2019-12-08 paper.rmd
revise intro

keywords          : "Latent Profile Analysis, mixture models, finite mixture models, tutorial, R, MPlus, mclust"
wordcount         : "4,882 (need to shorten to ~ 3,000)"

@@

For teachers, psychologists, researchers, and friends and family member, people
are highly-complex and not easily able to be characterized by a single
characteristic or personality trait. In the social sciences, broadly, and in
psychology, particularly, a statistical method that can be used to describe how
people, in their individual particularities, may have similarities on the basis
of some set of measures through which they can be grouped in meaningful,
distinctive ways. This approach, which has a provenance in developmental (or
*person-oriented*) approaches (Bergman & El-Khouri, 1997; Magnusson &
Cairns, 1996; see Linnenbrink-Garcia & Wormington, 2017, for a recent review),
though is now widely used in educational, social, personality, and other domains
of psychology, is an example of a general mixture model (Harring & Hodis, 2016;
Pastor, Barron, Miller, & Davis, 2007).

>>>>

keywords          : "Latent Profile Analysis, mixture models, finite mixture models, tutorial, R, MPlus, mclust"
wordcount         : ""

@@

In statistics classes, textbooks, and workshops, an example like the following is common:

> Grades are normally distributed, with $\mu$ = 75, $\sigma$ = 5.

Students, may (reasonably) ask, though, are grades *really* normally distributed?

```{r, include = FALSE}
library(simstudy)

def <- defData(varname = "x1", formula = 65, variance = 30)
def <- defData(def, varname = "x2", formula = 85, variance = 23)
def <- defData(def, varname = "xMix", 
               formula = "x1 | .5 + x2 | .5", 
               dist = "mixture")

set.seed(2716)
dx <- genData(1000, def)
```

Are their one distribution?

```{r, echo - FALSe}
library(ggplot2)
ggplot(dx, aes(x = xMix)) +
  geom_density() +
  papaja::theme_apa() +
  ylim(0, .10)
```

Or two?

```{r, echo - FALSe}
library(tidyverse)
dx %>% 
  gather(key, val, x1:x2) %>% 
  ggplot(aes(x = val, fill = key)) +
  geom_density(alpha = .3) +
  papaja::theme_apa() +
  ylim(0, .10)
```

This kind of distribution (that is *bimodal*) is not exclusive to grades. For teachers, psychologists, researchers, and friends and family members, people
are highly-complex and not easily able to be characterized by *one*
characteristic or personality trait--and its distribution. In the social sciences, broadly, and in psychology, particularly, a statistical method that can be used to describe how
people, in their individual particularities, may have similarities on the basis
of some set of measures through which they can be grouped in meaningful,
distinctive ways. This approach, which has a provenance in developmental (or
*person-oriented*) approaches (Bergman & El-Khouri, 1997; Magnusson &
Cairns, 1996; see Linnenbrink-Garcia & Wormington, 2017, for a recent review) is
an example of a general mixture model (Harring & Hodis, 2016; Pastor, Barron, Miller, & Davis, 2007).
